---

title: "Untitled"
author: "Julio Portella"
date: "6/3/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
## Importing R libraries

Importing the libraries

```{r cars}
library(dplyr)
library(caret)
library(Boruta)
library(randomForest)
library(glmnet)
library(randomForestExplainer)
library(rfFC)
library(ROCR)
library(precrec)
library(precrec)
library(ROSE)
library(ALEPlot)
```

## Loading the data

Load the data 
```{r}
library(dplyr)
source_CPET <- read.csv("C:/Users/julio/Projects/HealthInformatics/health-informatics-capstone/jp-post-capstone/data/8_min_data_updated_corrected.csv", header=TRUE, sep=",")
#source_CPET <-source_CPET %>% select(-c(X, PatientId,SessionId , sex, age,BMI,HasAnaerobicThresholdMean, PredictedMaxHR))
source_CPET <-source_CPET %>% select(-c(X, PatientId,SessionId , sex, age,BMI))
```
Finding more data
Scaling the data
```{r}
#dat2 <- source_CPET %>% mutate_at(-c("CardiacLim ", "PulmonaryLim","MuscleSkeletalLim"), ~(scale(.) %>% as.vector))
dat2 <- source_CPET %>% mutate_at(c('MaxVO2_EST',
       'MaxO2_EST', 'PredictedMaxHR', 'CardiacLim', 'PulmonaryLim',
       'MuscleSkeletalLim', 'Healthy', 'PeakHeartRate', 'MeanHeartRate',
       'MinHeartRate', 'StdHeartRate', 'LowestVE.VCO2', 'PeakVE.VCO2',
       'MeanVE.VCO2', 'StdVE.VCO2', 'PeakVO2Real', 'DiffPeakVO2', 'DiffPeakHR',
       'DiffPercentPeakVO2', 'DiffPercentPeakHR', 'MeanRER', 'PeakRER',
       'LowestRER', 'MeanVE', 'PeakVE', 'LowestVE', 'MeanRR', 'PeakRR',
       'LowestRR', 'MeanVO2', 'PeakVO2', 'LowestVO2', 'MeanVCO2', 'PeakVCO2',
       'LowestVCO2', 'HRvsVO2Slope', 'VEvsVCO2Slope', 'MeanO2Pulse',
       'MaxO2Pulse', 'MinO2Pulse', 'StdO2Pulse', 'O2PulseDiff',
       'O2PulsePercent', 'first_half_VO2Slope', 'first_half_HRSlope',
       'first_half_VCO2Slope', 'first_half_VESlope', 'first_half_RERSlope',
       'first_half_RRSlope', 'first_half_O2Slope', 'first_half_VEVCO2Slope',
       'first_half_VEVO2Slope', 'second_half_VO2Slope', 'second_half_HRSlope',
       'second_half_VCO2Slope', 'second_half_VESlope', 'second_half_RERSlope',
       'second_half_RRSlope', 'second_half_O2Slope', 'second_half_VEVCO2Slope',
       'second_half_VEVO2Slope', 'X15_to_85_VO2Slope', 'X15_to_85_HRSlope',
       'X15_to_85_VCO2Slope', 'X15_to_85_VESlope', 'X15_to_85_RERSlope',
       'X15_to_85_RRSlope', 'X15_to_85_O2Slope', 'X15_to_85_VEVCO2Slope',
       'X15_to_85_VEVO2Slope', 'VTTime', 'VO2atVT', 'PercentTimeAfterVT',
       'VO2vsPeakVO2atVT'), ~(scale(.) %>% as.vector))
```
Selecting the cardiac limitation data with the random split
```{r}
cardiac_lim_data <-dat2 %>% select(-c(PulmonaryLim,MuscleSkeletalLim,Healthy))
sample_size = floor(0.8*nrow(cardiac_lim_data))
set.seed(2345)
picked = sample(seq_len(nrow(cardiac_lim_data)),size = sample_size)
train_data =cardiac_lim_data[picked,]
test_data =cardiac_lim_data[-picked,]
```

Using boruta for feature selection FOR THE CARDIAC LIMITATION
```{r}
boruta.train <- Boruta(CardiacLim~., data = cardiac_lim_data, doTrace = 2)
```

```{r}
boruta.train
```


```{r}
plot(boruta.train, xlab = "", xaxt = "n", type='l')
lz<-lapply(1:ncol(boruta.train$ImpHistory),function(i)
boruta.train$ImpHistory[is.finite(boruta.train$ImpHistory[,i]),i])
names(lz) <- colnames(boruta.train$ImpHistory)
Labels <- sort(sapply(lz,median))
axis(side = 1,las=2,labels = names(Labels),
at = 1:ncol(boruta.train$ImpHistory), cex.axis = 0.5)
#text(cex = 0.9)
#text(x = 1:length(boruta.train), y = par("usr")[3] - 2.5, labels = names(Labels), srt = 35, cex = 0.8, xpd = NA)
```
Picking only the confirmed
```{r}
boruta.df.chosen <- attStats(boruta.train)
boruta.df.chosen %>% filter(decision == 'Confirmed') %>% arrange(desc(meanImp))
```

Picking tentative features into important features
```{r}
final.boruta <- TentativeRoughFix(boruta.train)
print(final.boruta)
getSelectedAttributes(final.boruta, withTentative = F)
```


```{r}
boruta.df <- attStats(final.boruta)
class(boruta.df)
print(boruta.df)
#starwars %>% filter(mass > mean(mass, na.rm = TRUE))
boruta.df %>% filter(decision != 'Rejected') %>% arrange(desc(meanImp))
boruta.chosen <- boruta.df %>% filter(decision != 'Rejected') %>% arrange(desc(meanImp))
```
## Pulmonary Limitation Model

Selecting the pulmonary limitation data with the random split
```{r}
pulmonary_lim_data <-dat2 %>% select(-c(CardiacLim,MuscleSkeletalLim,Healthy))
sample_size = floor(0.8*nrow(pulmonary_lim_data))
set.seed(2345)
picked = sample(seq_len(nrow(pulmonary_lim_data)),size = sample_size)
pulm_train_data =pulmonary_lim_data[picked,]
pulm_test_data =pulmonary_lim_data[-picked,]
```

Using boruta for feature selection FOR THE PULMONARY LIMITATION
```{r}
boruta.train <- Boruta(as.factor(PulmonaryLim)~., data = pulmonary_lim_data, doTrace = 2)
```

```{r}
plot(boruta.train, xlab = "", xaxt = "n")
lz<-lapply(1:ncol(boruta.train$ImpHistory),function(i)
boruta.train$ImpHistory[is.finite(boruta.train$ImpHistory[,i]),i])
names(lz) <- colnames(boruta.train$ImpHistory)
Labels <- sort(sapply(lz,median))
axis(side = 1,las=2,labels = names(Labels),
at = 1:ncol(boruta.train$ImpHistory), cex.axis = 0.5)
```

```{r}
boruta.df <- attStats(boruta.train)
class(boruta.df)
print(boruta.df)
#starwars %>% filter(mass > mean(mass, na.rm = TRUE))
boruta.df %>% filter(decision != 'Rejected') %>% arrange(desc(meanImp))

```
Picking tentative features into important features
```{r}
final.boruta <- TentativeRoughFix(boruta.train)
print(final.boruta)
getSelectedAttributes(final.boruta, withTentative = F)
```


```{r}
boruta.df <- attStats(final.boruta)
class(boruta.df)
print(boruta.df)
#starwars %>% filter(mass > mean(mass, na.rm = TRUE))
boruta.df %>% filter(decision != 'Rejected') %>% arrange(desc(meanImp))

```


```{r}
other_lim_data <-dat2 %>% select(-c(CardiacLim,PulmonaryLim,Healthy))
sample_size = floor(0.8*nrow(other_lim_data))
set.seed(2345)
picked = sample(seq_len(nrow(other_lim_data)),size = sample_size)
train_data =other_lim_data[picked,]
test_data =other_lim_data[-picked,]
```

Using boruta for feature selection FOR THE CARDIAC LIMITATION
```{r}
boruta.train <- Boruta(MuscleSkeletalLim~., data = other_lim_data, doTrace = 2)
```

```{r}
plot(boruta.train, xlab = "", xaxt = "n")
lz<-lapply(1:ncol(boruta.train$ImpHistory),function(i)
boruta.train$ImpHistory[is.finite(boruta.train$ImpHistory[,i]),i])
names(lz) <- colnames(boruta.train$ImpHistory)
Labels <- sort(sapply(lz,median))
axis(side = 1,las=2,labels = names(Labels),
at = 1:ncol(boruta.train$ImpHistory), cex.axis = 0.5)
```

```{r}
boruta.df <- attStats(boruta.train)
class(boruta.df)
print(boruta.df)
#starwars %>% filter(mass > mean(mass, na.rm = TRUE))
boruta.df %>% filter(decision != 'Rejected') %>% arrange(desc(meanImp))

```

```{r}
final.boruta <- TentativeRoughFix(boruta.train)
print(final.boruta)
getSelectedAttributes(final.boruta, withTentative = F)
```


```{r}
boruta.df <- attStats(final.boruta)
class(boruta.df)
print(boruta.df)
#starwars %>% filter(mass > mean(mass, na.rm = TRUE))
boruta.df %>% filter(decision != 'Rejected') %>% arrange(desc(meanImp))

```


## Muscle Skeletal Limitation Model


Selecting the pulmonary limitation data with the random split
```{r}
ms_lim_data <-dat2 %>% select(-c(CardiacLim,PulmonaryLim))
sample_size = floor(0.8*nrow(ms_lim_data))
set.seed(2345)
picked = sample(seq_len(nrow(ms_lim_data)),size = sample_size)
ms_train_data =ms_lim_data[picked,]
ms_test_data =ms_lim_data[-picked,]
```

Using boruta for feature selection FOR THE MUSCLE SKELETAL LIMITATION
```{r}
boruta.train <- Boruta(as.factor(Condition)~., data = ms_lim_data, doTrace = 2)
```


```{r}
plot(boruta.train, xlab = "", xaxt = "n")
lz<-lapply(1:ncol(boruta.train$ImpHistory),function(i)
boruta.train$ImpHistory[is.finite(boruta.train$ImpHistory[,i]),i])
names(lz) <- colnames(boruta.train$ImpHistory)
Labels <- sort(sapply(lz,median))
axis(side = 1,las=2,labels = names(Labels),
at = 1:ncol(boruta.train$ImpHistory), cex.axis = 0.5)
```
```{r}
boruta.df <- attStats(boruta.train)
class(boruta.df)
print(boruta.df)
#starwars %>% filter(mass > mean(mass, na.rm = TRUE))
boruta.df %>% filter(decision != 'Rejected') %>% arrange(desc(meanImp))

```

now that we have the selected features, let's use random forest
```{r}
#train_data =cardiac_lim_data[picked,]
#test_data =cardiac_lim_data[-picked,]
selected.attributes <- getSelectedAttributes(final.boruta, withTentative = F)
main_rf_classifier = randomForest(as.factor(MuscleSkeletalLim) ~ X75HRSlope+X75VO2Slope+LowestRER+PeakRER	+PeakVE+LowestVE+LowestVCO2+PeakVCO2+StdO2Pulse+X75VCO2Slope+PeakRR+MeanVCO2+LowestRR+MinHeartRate+PeakHeartRate+MeanRER+X75VESlope+MeanVE+PeakVO2+MeanO2Pulse+PeakVO2Real+LowestVO2+DiffPeakVO2+MeanVO2+MeanHeartRate, data=ms_train_data, ntree=500, mtry=2, importance=TRUE,classwt=c(40,100), keep.inbag=TRUE,replace=FALSE)
print(main_rf_classifier)
```

On the test dataset
```{r}
prediction_for_table<-predict(main_rf_classifier,ms_test_data)
table(observed=ms_test_data$MuscleSkeletalLim,predicted=prediction_for_table)
```
```{r}
boruta.df <- attStats(final.boruta)
class(boruta.df)
print(boruta.df)
#starwars %>% filter(mass > mean(mass, na.rm = TRUE))
boruta.df %>% filter(decision != 'Rejected') %>% arrange(desc(meanImp))

```



Doing the test with the new dataset

Load the data 
```{r}
source_CPET_2 <- read.csv("C:/Users/julio/Projects/HealthInformatics/health-informatics-capstone/jp-post-capstone/DigitalDataEnriched_updated.csv", header=TRUE, sep=",")
source_CPET_2 <-source_CPET_2 %>% select(-c(X, PatientId,SessionId , sex, age,BMI))
names(source_CPET_2)
```
Finding more data
Scaling the data
```{r}
source_CPET_2$PrimaryCondition <- source_CPET_2$PrimaryCardiacLim*4 + source_CPET_2$PrimaryPulmonaryLim*3 + source_CPET_2$OtherPrimaryLim*2+source_CPET_2$Healthy
source_CPET_2$PrimaryCondition <- as.factor(source_CPET_2$PrimaryCondition)
#dat2 <- source_CPET %>% mutate_at(-c("CardiacLim ", "PulmonaryLim","MuscleSkeletalLim"), ~(scale(.) %>% as.vector))
data_scaled <- source_CPET_2 %>% mutate_at(c("MaxVO2_EST","PeakHeartRate","MeanHeartRate","MinHeartRate","StdHeartRate","LowestVE.VCO2","PeakVE.VCO2","MeanVE.VCO2","StdVE.VCO2","PeakVO2Real","DiffPeakVO2","MeanRER","PeakRER","LowestRER","MeanVE","PeakVE","LowestVE","MeanRR","PeakRR","LowestRR","MeanVO2","PeakVO2","LowestVO2","HRvsVO2Slope","VEvsVCO2Slope","MeanO2Pulse","MaxO2Pulse","MinO2Pulse","StdO2Pulse","X75VO2Slope","X75HRSlope","X75VCO2Slope","X75VESlope","X75RERSlope","X75RRSlope","X75O2Slope","VO2atVT","VO2vsPeakVO2atVT","MeanVCO2","PeakVCO2","LowestVCO2"), ~(scale(.) %>% as.vector))
```
Selecting the cardiac limitation data with the random split
```{r}
lim_data <-data_scaled %>% select(-c(PrimaryPulmonaryLim,PrimaryCardiacLim,OtherPrimaryLim,Healthy))
sample_size = floor(0.8*nrow(lim_data))
set.seed(2345)
picked = sample(seq_len(nrow(lim_data)),size = sample_size)
train_data =lim_data[picked,]
test_data =lim_data[-picked,]
```

Using boruta in a single multilabel category
```{r}
boruta.train <- Boruta(PrimaryCondition~., data = train_data, doTrace = 2)
```

```{r}
boruta.train
```

```{r}
plot(boruta.train, xlab = "", xaxt = "n", type='l')
lz<-lapply(1:ncol(boruta.train$ImpHistory),function(i)
boruta.train$ImpHistory[is.finite(boruta.train$ImpHistory[,i]),i])
names(lz) <- colnames(boruta.train$ImpHistory)
Labels <- sort(sapply(lz,median))
axis(side = 1,las=2,labels = names(Labels),
at = 1:ncol(boruta.train$ImpHistory), cex.axis = 0.5)
#text(cex = 0.9)
#text(x = 1:length(boruta.train), y = par("usr")[3] - 2.5, labels = names(Labels), srt = 35, cex = 0.8, xpd = NA)
```

Picking tentative features into important features
```{r}
final.boruta <- TentativeRoughFix(boruta.train)
print(final.boruta)
getSelectedAttributes(final.boruta, withTentative = F)
```


```{r}
boruta.df <- attStats(final.boruta)
class(boruta.df)
print(boruta.df)
#starwars %>% filter(mass > mean(mass, na.rm = TRUE))
boruta.df %>% filter(decision == 'Rejected') %>% arrange(desc(meanImp))

```

```{r}
cond_rf_classifier = randomForest(PrimaryCondition ~. -MeanVE-LowestRER-PeakVE.VCO2-PeakRER-MinO2Pulse-MeanRR-VO2vsPeakVO2atVT-MaxVO2_EST-HRvsVO2Slope-HRvsVO2Slope-HasAnaerobicThresholdMean-Observation	, data=train_data, ntree=500, mtry=2, importance=TRUE,keep.inbag=TRUE,replace=FALSE)
prediction_for_table<-predict(cond_rf_classifier,test_data)
table(observed=test_data$PrimaryCondition,predicted=prediction_for_table)
```



```{r}
library(MASS)
# Fit the model
train_data$PrimaryCondition <- train_data$PrimaryCondition +1

model <- lda(PrimaryCondition~MeanHeartRate+PeakVO2Real+DiffPeakVO2+PeakRER+MeanVO2+PeakVO2+LowestVO2+HRvsVO2Slope+MeanO2Pulse+StdO2Pulse+X75VO2Slope+X75HRSlope+X75VCO2Slope+X75VESlope+X75RERSlope+X75RRSlope+X75O2Slope  , data = train_data)
# Make predictions
predictions <- model %>% predict(test_data)
# Model accuracy
mean(predictions$class==test_data$PrimaryCondition)
```

```{r}
library(MASS)
model <- lda(PrimaryCondition~~MeanHeartRate+PeakVO2Real+DiffPeakVO2+PeakRER+MeanVO2+PeakVO2+LowestVO2+HRvsVO2Slope+MeanO2Pulse+StdO2Pulse+X75VO2Slope+X75HRSlope+X75VCO2Slope+X75VESlope+X75RERSlope+X75RRSlope+X75O2Slope , data = train_data)
model
```






